---
title: "MSG analyses"
author: "Defu Yap, Geoffrey Brookshire, Daniel Casasanto"
output:
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE) # Don't print the R code
# knitr::opts_chunk$set(cache=FALSE)

source('stratified.R') # Requires data.table
library(knitr) # For pretty tables in the text (also pander or xtable)
library(parallel) # for parSapply
library(permute) # For within-subjects randomization: how() & shuffle()
library(analysisHelpers) # For fix.factors
library(ggplot2)
library(plotrix)

# Calculate the number of cores for parallel processing
ncores <- detectCores() - 1
# Initiate cluster
cl <- makeCluster(ncores)

```

# Introduction

Let's use this script to gather together all the analyses and to generate the plots.

# Method

## Data cleaning

Before all analysis, we excluded non-gestures and emblems. We also excluded gestures that were not produced in the four directions of the stories.

``` {r cleanData}

dat <- read.csv('data/msgclause.csv')
dat$Sub <- factor(paste(dat$Sub, dat$experiment, sep ='_')) # tag experiment to sub
names(dat)[names(dat) == 'direction'] <- 'gestureDirection' 
dat$Output.Language <- factor(paste(dat$Output.Language, "SL", sep ='')) # make same labels for both lang.types
dat$stroke.barrier <- ifelse(dat$stroke.barrier == "Y", "Y", "N") # cleaning up gestures above barrier coding
dat$stroke.barrier <- factor(dat$stroke.barrier)
# descriptives of gestures produced above the barrier
# addmargins(table(dat$experiment,dat$storyGesture, dat$stroke.barrier))
# round(addmargins(prop.table(table(dat$experiment, dat$storyGesture,dat$stroke.barrier)))*100,2)

# # descriptives of all gestures
# addmargins(table(dat$gestureDirection, dat$storyGesture, dat$Target.NonTarget))
# round(addmargins(prop.table(table(dat$gestureDirection, dat$storyGesture, dat$Target.NonTarget)))*100,2)

# # descriptives of analyzed getures
# s <- dat
# s$storyGesture <- factor(s$storyGesture, levels=c('BEAT','METAPHORIC','ICONIC','DEICTIC')) # exclude adaptors & emblems
# s$gestureDirection <- factor(s$gestureDirection, levels=c('UP','DOWN','LEFT','RIGHT')) # exclude directions not of interest
# addmargins(table(s$gestureDirection, s$storyGesture, s$Target.NonTarget))
# round(addmargins(prop.table(table(s$gestureDirection, s$storyGesture, s$Target.NonTarget)))*100,2)

#############################################
# Flow (all gestures, then beat gestures only)
  # 1. Lexically-affiliated gestures
    # i. Clause-wise congruity - Target clasuses only ('Output.Language' & 'ClauseDirection')
  # 2. Conceptually-affiliated gestures
    # i. Story-wise congruity - All clauses ('Input.Lang' & 'StoryDirection')   
    # ii. Story-wise congruity - All clauses ('Input.Lang' & 'StoryDirection')

# 1. subset for lexically-affilated 
dat <- subset(dat, Target.NonTarget == 'Target') # target clauses only
dat$Output.Language <- factor(dat$Output.Language) # drop Non-target clauses (dat$Output.Language == 'SL')
dat$ClauseDirection <- factor(tolower(dat$ClauseDirection))

# 2a. subset for conceptually-affiliated gestures
# dat <- subset(dat, Target.NonTarget == 'NonTarget') # non-target clauses only
# lang.type <- 'Input.Lang' # 'Input.Lang' /'Output.Language'
# direction.var <- 'StoryDirection' # StoryDirection' / 'ClauseDirection'
# clause.type <- 'Target.NonTarget'

# 2b. subset for lexically-affiliated gestures
lang.type <- 'Output.Language'
direction.var <- 'ClauseDirection'
clause.type <- 'Target.NonTarget'

dat <- dat[dat[[direction.var]] %in% c('up','down','left','right'),] # select four clause directions of interests
dat[[direction.var]] <- factor(dat[[direction.var]]) # drop unused levels

# exclude non-gestures and emblems
msg <- dat[which(dat$storyGesture != "" 
                 & dat$storyGesture != "ADAPTOR"
                 & dat$storyGesture != 'EMBLEM'
                 & dat$storyGesture != "NA "),] 
msg$storyGesture <- factor(msg$storyGesture)

# include only gestures produced in four directions of the stories
msg <- subset(msg, gestureDirection %in% c('UP','DOWN','LEFT','RIGHT'))
msg$gestureDirection <- factor(tolower(msg$gestureDirection)) 
msg$gestureDirection <- factor(msg$gestureDirection, levels = c('right','up','left','down'))

# select gesture-subset of interest
msg <- subset(msg, storyGesture != 'BEAT') # beat gestures only

```

## Analysis

We computed congruity as the proportion of gestures that were in the same direction as the dominant direction of the story.

``` {r congruityFunc}

# Calculate the proportion of congruent gestures in a data.frame
Congruity <- function(dat) {
  mean(dat[[direction.var]] == dat$gestureDirection)
}

# Function to help report the results.
# Makes a string reporting the stats for this test.
# Args
# emp: an empirical value
# perm.dist: a permuted distribution
ReportPermute <- function(emp, perm.dist) {
  pval <- mean(perm.dist > emp)
  paste('(', round(mean(perm.dist), 3) * 100,
      ' $\\pm$ ', round(sd(perm.dist), 3) * 100,
      '%, p = ', pval, ')',
      sep='')
}

```

### Overall congruity

Since there were four possible directions of interest, it would be tempting to treat 25\% as chance. However,  participants produced unequal numbers of gestures in each direction and each story type, therefore we cannot compare our data against an a priori chance value. To test whether gestures were congruent with the story direction more often than would be expected by chance, we performed within-subjects permutation tests. In each permutation, we randomly shuffled the story direction within each participant's data 10,000 times, and then computed the congruity over the shuffled data. The permuted distributions are reported as Mean $\pm$ SD.

``` {r overallPermuteFunc}
clusterExport(cl, c('shuffle', 'Congruity', 'how'))

# Within-subjects permutation test on proportion of congr. gestures.
WithinSubsCongPermute <- function(dat, n.reps=10000){
  CTRL <- how(blocks = dat$Sub) 
  PermHelper <- function() {
    shuff.inx <- shuffle(nrow(dat), control = CTRL) # shuffle indices within Sub
    dat[[direction.var]] <- dat[[direction.var]][shuff.inx]
    return(Congruity(dat)) # calculate mean congruence across all Sub
  }
  parSapply(cl, 1:n.reps, function(x) PermHelper()) # Parallel
  # replicate(n.reps, expr = PermHelper())
}

```

### Differences in congruity between visibility conditions

We performed a permutation test to determine whether the mean proportion of congruity differs between the different visibility conditions. We computed the variance of the proportion congruity between the three groups, and then randomly shuffled the visibility (experiment) label for each trial ($k = 10,000$).

```{r onewayAnovaPermuteFunc}
# Oneway permutation functions

# Compute the means for each input language; get the variance of those means.
GroupVar <- function(dat, IV) {
  group.means <- by(dat, dat[[IV]], Congruity)
  return(var(group.means))
}

# Permutation test analogous to a one-way ANOVA.
# Are the differences in congruity between groups greater than chance?
# F.K.A. WithinSubsCongExpVarPermute
OnewayPermTest <- function(dat, IV, n.reps=10000){
  PermHelper <- function() {
    CTRL <- how(blocks = dat$Sub)
    shuff.inx <- shuffle(nrow(dat), control = CTRL) # shuffle indices within Sub
    dat[[IV]] <- dat[[IV]][shuff.inx]
    return(GroupVar(dat, IV)) # calculate mean congruence across all Sub
  }
  parSapply(cl, 1:n.reps, function(x) PermHelper()) # Parallel
}

# Sort the data-frame by subject ahead of time (make sense)
msg <- msg[order(msg$Sub),]
OnewayPermTestBetween <- function(dat, n.reps=10000){
  PermHelper <- function() {
    # Which exp was each ppt in?
    subj.exp <- aggregate(experiment ~ Sub, data = msg, FUN = unique)
    # How many observations for each participant?
    n.obs.per.subj <- aggregate(experiment ~ Sub, data = msg, FUN = length)
    # Shuffle the experiments for each participant
    subj.exp$experiment <- subj.exp$experiment[shuffle(nrow(subj.exp))]
    dat$experiment <- rep(subj.exp$experiment, n.obs.per.subj$experiment) 
    return(GroupVar(dat, 'experiment')) # calculate mean congruence across all Sub
  }
  parSapply(cl, 1:n.reps, function(x) PermHelper()) # Parallel
}

# Report variance in the means of congruity
ReportOnewayPerm <- function(emp, perm.dist) {
  pval <- mean(perm.dist > emp)
  paste('( M = ', round(mean(perm.dist), 5) * 100,
      ' $\\pm$ ', round(sd(perm.dist), 5) * 100,
      '%, p = ', round(pval,3), ')',
      sep='')
}

```

We performed three separate permutation tests (k=10,000, each) to compare the pairwise differences in gesture congruity between each pair of visibility conditions. We randomly permuted the visibility conditions between each participant, and computed the pairwise differences in congruity. 

``` {r pairwisePermuteFunc}

# Compute the pairwise difference in congruity between groups.
PairwiseDiffHelper <- function(dat, IV, name1, name2) {
  c1 <- Congruity(dat[dat[[IV]] == name1,])
  c2 <- Congruity(dat[dat[[IV]] == name2,])
  return(c1 - c2)
}

BetweenSubsCongPairwisePermute <- function(dat, name1,name2, n.reps=10000){
  PermHelper <- function() {
    # Which exp was each ppt in?
    subj.exp <- aggregate(experiment ~ Sub, data = msg, FUN = unique)
    # How many observations for each participant?
    n.obs.per.subj <- aggregate(experiment ~ Sub, data = msg, FUN = length)
    # Shuffle the experiments for each participant
    subj.exp$experiment <- subj.exp$experiment[shuffle(nrow(subj.exp))]
    dat$experiment <- rep(subj.exp$experiment, n.obs.per.subj$experiment) 
    return(PairwiseDiffHelper(dat, 'experiment',name1,name2)) # calculate mean congruence across all Sub
  }
  parSapply(cl, 1:n.reps, function(x) PermHelper()) # Parallel
}

# Within-subjects permutation test on pairwise differences.
WithinSubsCongPairwisePermute <- function(dat, IV, name1, name2, n.reps=10000){
  PermHelper <- function() {
    CTRL <- how(blocks = dat$Sub)
    shuff.inx <- shuffle(nrow(dat), control = CTRL) # shuffle indices within Sub
    dat[[IV]] <- dat[[IV]][shuff.inx]
    return(PairwiseDiffHelper(dat, IV, name1, name2))
  }
  parSapply(cl, 1:n.reps, function(x) PermHelper())
}

# Within-subjects permutation test on pairwise differences between clause types.
# Target clauses: direction.var == Clausedirection and lang.type==output.language
# Non-Target clauses: direction.var==storydirection and lang.type==input.lang
# WithinSubsCongPairwisePermute <- function(dat, IV, name1, name2, n.reps=10000){
#   PermHelper <- function() {
#     CTRL <- how(blocks = dat$Sub)
#     shuff.inx <- shuffle(nrow(dat), control = CTRL) # shuffle indices within Sub
#     dat[[IV]] <- dat[[IV]][shuff.inx]
#     return(PairwiseDiffHelper(dat, IV, name1, name2))
#   }
#   parSapply(cl, 1:n.reps, function(x) PermHelper())
# }

```

We performed within-subjects permutations ($k=10,000$) to test whether congruity within each visibility condition was greater than would be expected by chance. 

### Congruity within each story direction

We performed bootstrapping tests ($k=1000$) to test whether congruity for each story direction was greater than would be expected by chance. To determine the chance level of congruity, we sampled gestures with replacement within each subject but across all story directions. To ensure that each subject had the same influence over the final congruity estimate, we sampled the same number of gestures as each subject produced in that story direction. We then compared the observed rate of congruity between the observed and predicted gesture directions against the chance level, established by bootstrapping. Effect sizes were reported in odds ratios and their confidence intervals.
 
``` {r singleDirectionBootFunc}
clusterExport(cl, c('stratifiedDT', 'setnames', 'setkeyv',
                    'is.data.table', 'as.data.table', 'data.table'))

# Bootstrap test on proportion of congruent gestures in one story direction.
BootSingleDirection <- function(dat, story.direction, n.reps=1000) {
  
  # First figure out how many samples to draw for each subject
  n.samps.per.subj <- aggregate(X ~ Sub, data=dat[dat[[direction.var]] == story.direction,],
                                FUN=length)
  counts <- n.samps.per.subj$X
  names(counts) <- n.samps.per.subj$Sub
  
  BootSampHelper <- function() {
    # Sample with replacement from within each group
    # Take the same number of observations as there 
    # are for each story direction.
    # Be sure to only look at subjects who had > 0 gestures in this cond.
    resamp.dat <- stratifiedDT(subset(dat, Sub %in% n.samps.per.subj$Sub),
                               group = 'Sub',
                               size = counts,
                               replace = TRUE)
    
    # Get the bootstrapped congruence with this story direction
    resamp.dat[[direction.var]] <- story.direction
    c.resamp <- Congruity(resamp.dat)
    return(c.resamp)
  }
  
  # replicate(n=n.reps, expr=BootSampHelper())
  parSapply(cl, 1:n.reps, function(x) BootSampHelper())
  
  # return mean congruity within each story direction
  
}

# Report the bootstrap results.
ReportBootstrap <- function(emp, c.resamp) {
  pval <- mean(c.resamp > emp)
  paste('(', round(mean(c.resamp), 3) * 100,
      ' $\\pm$ ', round(sd(c.resamp), 3) * 100,
      '%, p = ', pval, ')',
      sep='')
}

# Odds ratio of selected gesture and story direction.
ReportOddsRatio <- function(data, direc) {
  sel.gest.dir <- data$gestureDirection == direc # Selected gesture direction
  sel.story.dir <- data[[direction.var]] == direc # Selected story direction
  x <- table(sel.gest.dir, sel.story.dir)
  or <- fisher.test(x)
  paste('OR = ', round(or$estimate, 2),
        ', p = ', round(or$p.value,6),
        ', 95% CI [', round(or$conf.int[1],2),  
        ', ', round(or$conf.int[2],2), ']',
        sep='')
}

```

### Differences in congruity between language types

We performed a permutation test to determine whether the mean proportion of congruity differs between the different types of language. We computed the variance of the proportion congruity between the three groups, and then randomly shuffled the language label for each trial ($k = 10,000$).

We performed three separate permutation tests (k=10,000, each) to compare the pairwise differences in gesture congruity between each pair of language types. We randomly permuted the story directions within each participant, and computed the pairwise differences in congruity.

We performed within-subjects permutations ($k=10,000$) to test whether congruity within each language type was greater than would be expected by chance. 

### Differences in congruity with respect to the presence of directional meaning in clauses. 

We performed a permutation test (k=10,000) to compare the pairwise differences in gesture congruity between each pair of clause types. We randomly permuted the story directions within each participant, and computed the pairwise differences in congruity. Then we performed within-subjects permutations tests to test whether congruity within each type of clause (Target vs. Non-target clause) was greater than would be expected by chance ($k=10,000$).

# Results (Combined data)

## Descriptive statistics

``` {r allGestureDescriptives}
clusterExport(cl, c('direction.var', 'BootSingleDirection'))

# Counts and proportions of all types of gestures
#addmargins(table(dat$storyGesture)) # check gest counts
#round(prop.table(table(dat$storyGesture)),4) # check gestproportions
counts <- summary(dat$storyGesture)[1:6]
bar.names <- names(counts)
b1 <- barplot(height = counts,
        names.arg = bar.names,
        las = 2,
        col = 'blue',
        cex.names = 0.8, ylim = c(0,max(counts)+500),
        main = 'Total number of gestures produced',
        ylab = 'Count')
text(x = b1, y = counts, label = counts, pos = 3, cex = 0.8)
rm(counts, bar.names)

# Counts and proportions of gestures in all directions
counts <- summary(dat$gestureDirection)[1:11] # exclude NAs
bar.names <- names(counts)
b2 <- barplot(height = counts,
        names.arg = bar.names,
        las = 2,
        col = 'blue',
        cex.names = 0.8, ylim = c(0,max(counts)+500),
        main = 'Total number of gestures produced',
        ylab = 'Count')
## Add text at top of bars
text(x = b2, y = counts, label = counts, pos = 3, cex = 0.8)
rm(counts, bar.names)

# plot gesture counts according to gesture direction
bartable = table(msg$gestureDirection, msg$storyGesture)  
b3 <- barplot(bartable, beside = TRUE, 
              col = c('blue','red','brown','orange'),
              main = 'Total number of gestures produced',
              ylim = c(0,max(bartable)+500), ylab = 'Count',
              legend = levels(unique(msg$gestureDirection))) 
## Add text at top of bars
text(x = b3, y = bartable, label = bartable, pos = 3, cex = 0.8)

# barplot by proportions
bartable = prop.table(table(msg$gestureDirection, msg[[direction.var]]),2)  
b4 <- barplot(bartable, beside = TRUE, 
              col = c('blue','red','brown','orange'),
              main = 'Proportions of gestures produced',
              ylim = c(0,1), ylab = 'Proportions of gestures',
             xlab = 'Story directions',
              legend = levels(unique(msg$gestureDirection)),
             args.legend = list(title='gesture\ndirections', bty = 'n'))
## Add text at top of bars
text(x = b4, y = bartable, label = round(bartable,2), pos = 3, cex = 0.8)

# Calculate empirical and bootstrapped congruity in each language type and storydirection 
summ.perm <- by(msg, msg[[lang.type]], FUN = function(d) {
  dirs <- levels(d[[direction.var]])
  bs <- lapply(dirs, FUN = function(direc) {
    list(perm=BootSingleDirection(d, direc, n.reps=100), # Bootstrapped chance cong
         emp=Congruity(d[d[[direction.var]] == direc,])) # Empirical cong
    })
  names(bs) <- dirs # Make sure we know which story direction is which
  return(bs)
})

#Init plot
plot(NULL, NULL, ylim = c(0,1), xlim = c(1, 14),
     xlab='Language type', xaxt='n', ylab='Congruity')

box.pos <- c(1:4, 6:9, 11:14) # Keep track of where the box will be
i.box <- 1

axis(side = 1, labels = levels(msg[[lang.type]]),
     at = c(2.5, 7.5, 12.5), # line = 2,
     lty=0)

# 'Dictionary' of arrows for their labels
list.arrows <- list(left ='←', right = '→', up = '↑', down = '↓')
# text(box.pos, 0.05, rep(uni.arrows, times=3), cex=1.5)
msg[[direction.var]] <- factor(msg[[direction.var]], levels = c('up','down','left','right'))
text(box.pos, 0.05, rep(list.arrows[levels(msg[[direction.var]])], times=3), cex=1.5)
text(box.pos,adj=-.15, "Congruity in the four story directions within each language type",3)
for (languageType in levels(msg[[lang.type]])) {
  for (direction in levels(msg[[direction.var]])) {
    boxplot(summ.perm[[languageType]][[direction]]$perm,
            range = 0,
            # boxwex = 0.9,
            staplewex = 0.1, lty = 1,
            add = TRUE, at = box.pos[i.box], yaxt='n')
    points(box.pos[i.box], summ.perm[[languageType]][[direction]]$emp,
           pch=19, col='red')
    i.box <- i.box + 1
  }
}

```

### Overall congruity

``` {r overallChanceCongruity, cache=FALSE}

perm.stat <- WithinSubsCongPermute(msg) 

# If we want permutation histograms, here's how we'd do them
# hist(perm.stat, xlim=c(0,1), col = 'black')
# abline(v=Congruity(no.barrier), col='red')

```

- Empirical congruity across all story types and directions:  `r round(Congruity(msg), 3) * 100` % of the time. 
- Chance congruity:  `r ReportPermute(Congruity(msg), perm.stat)`.


``` {r visibilityChanceCongrutiy, cache=FALSE} 
clusterExport(cl, c('msg','GroupVar', 'PairwiseDiffHelper'))

# Congruity across visibility conditions
empVar <-GroupVar(msg,'experiment') # empirical value
permVar.dist <- OnewayPermTestBetween(msg) # chance distribution

# Pairwise difference in chance congruity between visibility conditions
empExp1v3Pair <- PairwiseDiffHelper(msg,'experiment','MSG1','MSG3')
permExp1v3Pair <- BetweenSubsCongPairwisePermute(msg,'MSG1','MSG3') 

empExp1v4Pair <- PairwiseDiffHelper(msg,'experiment', 'MSG1','MSG4')
permExp1v4Pair <- BetweenSubsCongPairwisePermute(msg,'MSG1','MSG4')

empExp3v4Pair <- PairwiseDiffHelper(msg, 'experiment','MSG4','MSG3')
permExp3v4Pair <- BetweenSubsCongPairwisePermute(msg,'MSG4','MSG3') 

# Congruity within each visibility condition
empExp1 <- Congruity(subset(msg, experiment=='MSG1'))
permExp1.dist <- WithinSubsCongPermute(subset(msg,experiment == 'MSG1'))

empExp3 <- Congruity(subset(msg, experiment=='MSG3'))
permExp3.dist <- WithinSubsCongPermute(subset(msg,experiment == 'MSG3'))

empExp4 <- Congruity(subset(msg, experiment=='MSG4'))
permExp4.dist <- WithinSubsCongPermute(subset(msg,experiment == 'MSG4'))  

```

### Congruity across visibility conditions

In this section, we'll want to include the 'one-way ANOVA' across all three visibility conditions, as well as pairwise permutation tests. Also, include tests of whether congruity is greater than chance in each visibility condition.We randomly permuted the visibility conditions between each participant, and computed the variance of the mean congruity across visibility conditions. 

- The observed variance in congruity: `r round(empVar,4)*100`%.
- Permuted chance variance `r ReportOnewayPerm(empVar, permVar.dist)`.

We performed three separate permutation tests (k=10,000, each) between each pair of visibility conditions. We randomly permuted the story directions within each participant, and computed the pairwise differences in congruity. 

- No-barrier vs. barrier-only condition: `r round(empExp1v3Pair, 3) * 100`%, chance rate = `r ReportPermute(empExp1v3Pair,permExp1v3Pair)`.
- No-barrier vs. barrier-with-blindfold condition: `r round(empExp1v4Pair, 3) * 100`%, chance rate = `r ReportPermute(empExp1v4Pair,permExp1v4Pair)`.
- Barrier-only vs. barrier-with-blindfold condition: `r round(empExp3v4Pair, 3) * 100`%, chance rate = `r ReportPermute(empExp3v4Pair,permExp3v4Pair)`.

We performed within-subjects permutations (k=10,000) to test whether congruity within each visibility condition was greater than would be expected by chance. Participants produced congruent gestures more often than would be expected by chance in all three visibility conditions.

- no-barrier condition: `r round(empExp1, 3) * 100`% congruent  (chance rate = `r ReportPermute(empExp1, permExp1.dist)`); 
- barrier condition: `r round(empExp3, 3) * 100`% congruent (chance rate =  `r ReportPermute(empExp3, permExp3.dist)`) , and 
- barrier-and-blindfold condition: `r round(empExp4, 3) * 100`% congruent (chance rate =  `r ReportPermute(empExp4, permExp4.dist)`).

Since the overall congruity in all three visibility conditions are significantly above chance, we collapsed the visibility conditions for the rest of the analyses.

### Congruity in each direction

``` {r singleDirectionRadialPlots, fig.height=12, fig.width=12}

# Table of gesture-direction produced in each story direction
# t <- xtabs(~ StoryDirection + gestureDirection, data=msg) # original xtabs
# t <- xtabs(msg[c(direction.var, 'gestureDirection'),]) # xtabs failed but table works!
t <- table(msg[c(direction.var, 'gestureDirection')]) 
# t <- t[,c('right','up','left','down')] #rearrange coloumns for radial.plot 

# Radial plot function for gestures produced in each story direction
MyRadPlot <- function(x, ...) {
  radial.plot(x,
            labels=c('RIGHT','UP','LEFT','DOWN'),
            rp.type='p',
            poly.col=c(rgb(1,0,0,.5),rgb(0,1,0,.5)),
            line.col=c('black','black'),lwd=1,
            show.grid.labels = 3, boxed.radial = F, bty='n',
            ...)
}

# Radial plots of gestures in each story direction
par(mfrow=c(2,2))
# lapply(unique(), MyRadPlot, data = t )
MyRadPlot(t['down',], radial.lim=c(0, max(t)), main='Downward stories')
MyRadPlot(t['up',], radial.lim=c(0, max(t)), main='Upward stories')
MyRadPlot(t['left',], radial.lim=c(0, max(t)), main='Leftward stories')
MyRadPlot(t['right',], radial.lim=c(0, max(t)), main='Rightward stories')


# Radial representation of the same lang.type by direction.var boxplots
par(mfrow=c(5,5))

# # Function to plot radials for each story direction according to each language type
# for(languageType in levels(msg[[lang.type]])) {
#   for(direction in levels(msg[[direction.var]])) {
#     MyRadPlot(g[,,languageType][direction.var,],
#               radial.lim=c(0, max(g[,,languageType])), 
#               main=paste0(direction.var),'ward stories')
#   }
# }

g <- table(msg[c(direction.var, 'gestureDirection', lang.type)])
par(mar = c(0,0,0,0)) # LSL
plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, paste("LSL"), cex = 1.6, col = "black")
MyRadPlot(g[,,'LSL']['down',], radial.lim=c(0, max(g[,,'LSL'])), main='Downward stories')
MyRadPlot(g[,,'LSL']['up',], radial.lim=c(0, max(g[,,'LSL'])), main='Upward stories')
MyRadPlot(g[,,'LSL']['left',], radial.lim=c(0, max(g[,,'LSL'])), main='Leftward stories')
MyRadPlot(g[,,'LSL']['right',], radial.lim=c(0, max(g[,,'LSL'])), main='Rightward stories')

par(mar = c(0,0,0,0)) # MSL
plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, paste("MSL"), cex = 1.6, col = "black")
MyRadPlot(g[,,'MSL']['down',], radial.lim=c(0, max(g[,,'MSL'])), main='Downward stories')
MyRadPlot(g[,,'MSL']['up',], radial.lim=c(0, max(g[,,'MSL'])), main='Upward stories')
MyRadPlot(g[,,'MSL']['left',], radial.lim=c(0, max(g[,,'MSL'])), main='Leftward stories')
MyRadPlot(g[,,'MSL']['right',], radial.lim=c(0, max(g[,,'MSL'])), main='Rightward stories')

par(mar = c(0,0,0,0)) # NSL
plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, paste("NSL"), cex = 1.6, col = "black")
MyRadPlot(g[,,'NSL']['down',], radial.lim=c(0, max(g[,,'NSL'])), main='Downward stories')
MyRadPlot(g[,,'NSL']['up',], radial.lim=c(0, max(g[,,'NSL'])), main='Upward stories')
MyRadPlot(g[,,'NSL']['left',], radial.lim=c(0, max(g[,,'NSL'])), main='Leftward stories')
MyRadPlot(g[,,'NSL']['right',], radial.lim=c(0, max(g[,,'NSL'])), main='Rightward stories')

c <- table(msg[c(direction.var, 'gestureDirection', 'Target.NonTarget')])
par(mar = c(0,0,0,0)) # Target clauses
plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, paste("Target\nclauses"), cex = 1.6, col = "black")
MyRadPlot(c[,,'Target']['down',], radial.lim=c(0, max(c[,,'Target'])), main='Downward stories')
MyRadPlot(c[,,'Target']['up',], radial.lim=c(0, max(c[,,'Target'])), main='Upward stories')
MyRadPlot(c[,,'Target']['left',], radial.lim=c(0, max(c[,,'Target'])), main='Leftward stories')
MyRadPlot(c[,,'Target']['right',], radial.lim=c(0, max(c[,,'Target'])), main='Rightward stories')

par(mar = c(0,0,0,0)) # Non-target clauses
plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
text(x = 0.5, y = 0.5, paste("Non-target\nclauses"), cex = 1.6, col = "black")
MyRadPlot(c[,,'NonTarget']['down',], radial.lim=c(0, max(c[,,'NonTarget'])), main='Downward stories')
MyRadPlot(c[,,'NonTarget']['up',], radial.lim=c(0, max(c[,,'NonTarget'])), main='Upward stories')
MyRadPlot(c[,,'NonTarget']['left',], radial.lim=c(0, max(c[,,'NonTarget'])), main='Leftward stories')
MyRadPlot(c[,,'NonTarget']['right',], radial.lim=c(0, max(c[,,'NonTarget'])), main='Rightward stories')

```

```{r singleDirectionChanceBootstrappedRates, cache=FALSE}
clusterExport(cl, 'direction.var')

emp.up <- Congruity(msg[msg[[direction.var]] == 'up',]) # emp congruity
boot.up <- BootSingleDirection(msg,'up') # chance distribution

emp.down <- Congruity(msg[msg[[direction.var]] == 'down',])
boot.down <- BootSingleDirection(msg,'down')

emp.left <- Congruity(msg[msg[[direction.var]] == 'left',])
boot.left <- BootSingleDirection(msg,'left')

emp.right <- Congruity(msg[msg[[direction.var]] == 'right',])
boot.right <- BootSingleDirection(msg,'right')

```

In all four story directions, participants gestured congruently with the story direction more often than would be expected by chance. 

- Upward stories had a `r round(emp.up, 3) * 100`% observed rate of congruity (chance rate = `r ReportBootstrap(emp.up,boot.up)`%, `r ReportOddsRatio(msg,'up')`. 
- Downward stories had a `r round(emp.down, 3) * 100`% observed rate of congruity (chance rate = `r ReportBootstrap(emp.down,boot.down)`%, `r ReportOddsRatio(msg,'down')`. 
- Leftward stories had a `r round(emp.left, 3) * 100`% observed rate of congruity (chance rate = `r ReportBootstrap(emp.left,boot.left)`%, `r ReportOddsRatio(msg,'left')`. 
- Rightward stories had a `r round(emp.right, 3) * 100`% observed rate of congruity (chance rate = `r ReportBootstrap(emp.right,boot.right)`%, `r ReportOddsRatio(msg,'right')`.

### Congruity across literal, metaphorical, and non-spatial stories

``` {r OneWayPermLanguageTypes, cache=FALSE} 
clusterExport(cl, 'lang.type')

permVar.dist <- OnewayPermTest(msg, direction.var)

```

In this section, we'll want to include the 'one-way ANOVA' across all three language types, as well as pairwise permutation tests. Also, include tests of whether congruity is greater than chance in each language type.

We randomly permuted the language types within each participant, and computed the variance of the mean congruity across language types. 

- Observed variance: `r round(GroupVar(msg,direction.var),4)*100`%
- Permuted chance variance: `r ReportOnewayPerm(GroupVar(msg,direction.var), permVar.dist)`.

``` {r pairwisediffChanceCongrutiyLanguageTypes, cache=FALSE}
emp.MLPair <- PairwiseDiffHelper(msg,lang.type,'MSL','LSL')
perm.MLPair <- WithinSubsCongPairwisePermute(msg,lang.type,'MSL','LSL') 

emp.NLPair <- PairwiseDiffHelper(msg,lang.type,'NSL','LSL')
perm.NLPair <- WithinSubsCongPairwisePermute(msg,lang.type,'NSL','LSL')

emp.MNPair <- PairwiseDiffHelper(msg,lang.type,'MSL','NSL')
perm.MNPair <- WithinSubsCongPairwisePermute(msg,lang.type,'MSL','NSL') 

```

### Pairwise differences in congruity between language types
We performed three separate permutation tests (k=10,000, each) between each pair of language types. We randomly permuted the story directions within each participant, and computed the pairwise differences in congruity. 

- LSL vs. MSL (`r round(emp.MLPair, 3)`, chance rate = `r ReportPermute(emp.MLPair,perm.MLPair)`
- NSL vs. LSL (`r round(emp.NLPair,3)`, chance rate = `r ReportPermute(emp.NLPair,perm.NLPair)`
- NSL vs. MSL (`r round(emp.MNPair,3)`, chance rate = `r ReportPermute(emp.MNPair,perm.MNPair)`.

``` {r chanceCongruityLanguageTypes, cache=FALSE}

emp.LSL <- Congruity(msg[msg[[lang.type]] == 'LSL',]) # emp congruity
permLSL.dist <- WithinSubsCongPermute(msg[msg[[lang.type]] == 'LSL',])

emp.MSL <- Congruity(msg[msg[[lang.type]] == 'MSL',]) 
permMSL.dist <- WithinSubsCongPermute(msg[msg[[lang.type]] == 'MSL',])

emp.NSL <- Congruity(msg[msg[[lang.type]] == 'NSL',]) 
permNSL.dist <- WithinSubsCongPermute(msg[msg[[lang.type]] == 'NSL',])

```

### Congruity in each language type
We performed within-subjects permutations (k=10,000) to test whether congruity within each language type was greater than would be expected by chance. Participants produced congruent gestures more often than would be expected by chance in all three language types.

- LSL: `r round(emp.LSL, 3) * 100`% congruent  (chance rate = `r ReportPermute(emp.LSL, permLSL.dist)`); 
- MSL: `r round(emp.MSL, 3) * 100`% congruent (chance rate =  `r ReportPermute(emp.MSL, permMSL.dist)`), 
- NSL: `r round(emp.NSL, 3) * 100`% congruent (chance rate =  `r ReportPermute(emp.NSL, permNSL.dist)`).

``` {r pairwisediffChanceCongrutiyClauseTypes, cache=FALSE}
# this is wrong. Check with Geoff again:
# congruity in target clauses based on story direction
# congruity in non-target clauses based on clause direction

clusterExport(cl, 'clause.type')

emp.TNTPair <- PairwiseDiffHelper(msg,clause.type,'Target','NonTarget')
perm.TNTPair <- WithinSubsCongPairwisePermute(msg,clause.type,'Target','NonTarget') 

```

### Pairwise differences in congruity between clause types
We performed three separate permutation tests (k=10,000, each) between clause types (target vs. non-target). We randomly permuted the story directions within each participant, and computed the pairwise differences in congruity. 

- Target vs. Non-target clauses (`r round(emp.TNTPair, 3)`, chance rate = `r ReportPermute(emp.MLPair,perm.TNTPair)`

``` {r chanceCongruityClauseTypes, cache=FALSE}
# # crap! Target clauses (lexical affiliates are based on clause level) BUT non-targe clauses are based on story level. How do we account for the different level of analyses?
# 
# # story-level congruity - so not change 'lang.type' & 'direction.var'
# nontarg.clauses <- subset(msg, Target.NonTarget == 'NonTarget') # non-target clauses only
# emp.NonTarget <- Congruity(nontarg.clauses[nontarg.clauses$Target.NonTarget == 'NonTarget',]) 
# permNonTarget.dist <- WithinSubsCongPermute(nontarg.clauses[nontarg.clauses$Target.NonTarget == 'NonTarget',])
# 
# # clause-level congruity (based on clause direction) - lexical affiliates (redo Rmarkdown)
# targ.clauses <- subset(msg, Target.NonTarget == 'Target') # non-target clauses only
# lang.type <- 'Output.Language'
# direction.var <- 'ClauseDirection' 
# targ.clauses <- targ.clauses[targ.clauses[[direction.var]] %in% c('up','down','left','right'),] # select four clause directions of interests
# targ.clauses[[direction.var]] <- factor(targ.clauses[[direction.var]]) # drop unused levels
# emp.Target <- Congruity(targ.clauses) # emp congruity
# permTarget.dist <- WithinSubsCongPermute(targ.clauses)
# table(targ.clauses$ClauseDirection, targ.clauses$gestureDirection)


### Congruity in each clause type (collapsing other factors)
# We performed within-subjects permutations (k=10,000) to test whether congruity within each language type was greater than would be expected by chance. Participants produced congruent gestures more often than would be expected by chance in all three language types. -->

# Target clauses: `r round(emp.Target, 3) * 100`% congruent  (chance rate = `r ReportPermute(emp.Target, permTarget.dist)`); 
# Non-Target clauses: `r round(emp.NonTarget, 3) * 100`% congruent (chance rate =  `r ReportPermute(emp.NonTarget, permNonTarget.dist)`).
```
